Wealth dynamics in the presence of network structure and primitive cooperation

### 2019-07-12 manuscript/analysis.tex
Add life expectancy section.

\begin{figure}[hb]
\centering
\includegraphics[width=\columnwidth]{figures/giniVsLambda.png}
\caption{The average Gini coefficient, computed before the onset of starvation,
for various values of the ER $\lambda$ connectivity parameter.}
\label{fig:giniVsLambda}
\end{figure}

@@

% Plots from 7/12 showing you best get in a proto if lambda and white noise 
%   are high

>>>>

\begin{figure}[hb]
\centering
\includegraphics[width=\columnwidth]{figures/giniVsLambda.png}
\caption{The average Gini coefficient, computed before the onset of starvation,
for various values of the ER $\lambda$ connectivity parameter. The color band
represents a bootstrapped 95\% confidence interval.}
\label{fig:giniVsLambda}
\end{figure}

@@

Rather than absolute wealth, which the Gini coefficient measures, an alternate
measure of well-being is the ability to survive an economic downturn. This,
after all, is the chief benefit an agent should be able to expect from joining
a proto: it serves as a kind of insurance policy against future poverty. It is
therefore interesting to compare the life expectancy of agents who join protos
with those who do not.

There are many factors at play here, one of which is the level of white noise
($\sigma^2$) in the agents' income. Figure~\ref{fig:avgLifespanLambda} depicts
how the life expectancy of isolates and non-isolates depends on $\sigma^2$ for
two different values of $\lambda$. The top plot shows that for relatively
stable agent income levels, there is not much difference between the two lines
-- and hence, not much advantage (or disadvantage) to an agent's joining a
proto. Interestingly, however, the more volatile the income stream becomes, the
more benefit there is to pooling resources. The effect is even more pronounced
with more densely connected graphs, as in the bottom plot: here, when income is
more noisy, agents who join protos live nearly twice as long as those who
don't.

% Insert Rajesh's fascinating speculations about why this should be the case.

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{figures/avgLifespanLambda_025.png}
\includegraphics[width=\columnwidth]{figures/avgLifespanLambda2.png}
\caption{Life expectancy comparison between isolates (non-proto members) and
non-isolates (proto members) for different values of $\lambda$ and $\sigma^2$.
The \texttt{salary} parameter was set to 20, so the x-axis ranges from a nearly
constant agent income to a scenario when the noise is as high as the average.}
\label{fig:avgLifespanLambda}
\end{figure}

### 2019-07-12 manuscript/draftcssa19.tex
Add spot for Will's Gini param sweep plot.
Add life expectancy section.

% SD: In light of recent discussions, we may want to downplay Gini and up-play the robustness of agents to exogenous shocks.
The phenomenon under scrutiny is the emergence and evolution of economic inequality in societies with non-trivial social structures. The model we use to answer questions surrounding this phenomenon must possess a dynamic model of wealth accumulation, a model of social structure, and a suitably useful measure of inequality. The dynamics are modeled as Brownian noise driven linear dynamical systems, here constant growth rate; the structure is modeled by networks, here random graphs; the measures of inequality used are Gini coefficients, and survival time distributions of agents in response to lack of resources. Although both the dynamical system and the network model are quite well understood, the precise interplay of structure and dynamics produces interesting emergent wealth distributions. To the best of our knowledge, this specific combination of network structure and social dynamics has not been discussed in the computational social science (CSS) or mathematical sociology literature, and we consider this the primary contribution of this work.

@@

The simulation proceeds as follows. For configurable parameters
$N > 0$,
$\texttt{init\_max} > 0$,
$\texttt{salary} > 0$,
$m > 0$ (metabolic rate),
$t_p > 0$ (the ``proto threshold''),
$\sigma^2 \ge 0$ and
$0 \le \lambda \le 1$:

@@

@@

\section{Analysis}
%\input{analysis}

>>>>

\label{giniKindaSucks}
% SD: In light of recent discussions, we may want to downplay Gini and up-play the robustness of agents to exogenous shocks.
The phenomenon under scrutiny is the emergence and evolution of economic inequality in societies with non-trivial social structures. The model we use to answer questions surrounding this phenomenon must possess a dynamic model of wealth accumulation, a model of social structure, and a suitably useful measure of inequality. The dynamics are modeled as Brownian noise driven linear dynamical systems, here constant growth rate; the structure is modeled by networks, here random graphs; the measures of inequality used are Gini coefficients, and survival time distributions of agents in response to lack of resources. Although both the dynamical system and the network model are quite well understood, the precise interplay of structure and dynamics produces interesting emergent wealth distributions. To the best of our knowledge, this specific combination of network structure and social dynamics has not been discussed in the computational social science (CSS) or mathematical sociology literature, and we consider this the primary contribution of this work.

@@

The simulation proceeds as follows. For configurable parameters
$N > 0$,
$\texttt{init\_max} > 0$,
$\texttt{salary} > 0$,
$0 \le m \le salary$ (metabolic rate),
$\sigma^2 \ge 0$,
$t_p > 0$ (the ``proto threshold''), and
$0 \le \lambda \le 1$:

@@

Importantly, take note that in this simple model, every non-isolate agent (\textit{viz.}, every agent with at least one graph neighbor) will eventually join a proto. This is important in the analysis that follows, as we will compare isolates with non-isolates in an attempt to measure the benefits of proto membership.

@@

\section{Verification}
\input{verification}

\section{Analysis}
\input{analysis}
